---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tba

<!-- badges: start -->
<!-- badges: end -->

The goal of the tba package is to provide easy access to match data for events 
run in the first robotics division. Access is based on the Read API v3 of [The Blue Alliance](https://www.thebluealliance.com/). 

## Installation

You can install the development version of tba from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("heike/tba")
```

In order to access the API of The Blue Alliance, you will need to [sign up for a (free) account](https://www.thebluealliance.com/apidocs/v3). 
Then run 
``` r
tba::set_api_key(key = "jnn1ZiUfqPuA9UJyV9EVa3jtL5qudBBlYXKPEVQDeQgQgz5L8BypwjTzVKxtCs3W") 
# key is not valid, get your own!
```


## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(tba)
## basic example code

# Team Neutrino has team code `frc3928` 
get_records("team/frc3928")

# In the 2024 season, the team attended the following events:
events <- get_records("team/frc3928/events/2024")
events$name
events$event_code
```

### Access Score Data from a Competition

The matches from *Clash in the Corn* can be accessed using the event code `2024iawes`:

```{r message=FALSE}
library(dplyr)
matches <- get_records("event/2024iawes/matches")
count(matches, comp_level)
# 22 qualifying matches
# 5 semi-finals
# 2 finals
```

The function `get_match_details` turns the records into a more manageable form by turning the match-based rows from the `get_records` results into team-based results, i.e.  every row corresponds to the 
detailed scores of each team in each one of the matches:

```{r}
scores <- get_match_details(matches)
head(scores %>% select(comp_level, match_number, set_number, alliance, team_key, alliances_score, ends_with("Points")))
```

### Calculate a team's contribution to an outcome

The function `get_ranking_by` allows a ranking of each team (not the alliance!)  by any numeric variable. It only makes sense to use matches from the qualifying round. The resulting number is an estimate of the team's contribution to the chosen measurement. 
 When using `get_ranking_by` with the `score` from the qualifying matches, the resulting coefficients for the ranking is the Offesnive Power Rating (OPR) - as published on the TBA API under insights.

```{r}
# ranking by OPR
scores %>% filter(comp_level=="qm") %>%
  get_ranking_by(alliances_score)
```

It makes sense, to not include points from fouls (assuming that it is hard to make another team foul one's alliance):

```{r}
no_fouls <- scores %>% filter(comp_level == "qm") %>%
  get_ranking_by(alliances_score - score_breakdown_foulPoints)
head(no_fouls)
```

Most of the time, the resulting ranking should be quite similar, but differences in these rankings might give some insight as to whether some teams profitted from being fouled during the qualification round. In this example, most of the rankings stay the same, but the 2nd and the 3rd team switch places:  `frc5041` seems to have benefitted during the qualifying matches more from their opponents' fouls than `frc4646`. Their rating of contribution to the team's score without counting fouls is quite similar. 

```{r}
scores %>% filter(comp_level=="qm") %>%
  get_ranking_by(alliances_score-score_breakdown_foulPoints)
```

### Comparing Multiple Outcomes

By comparing the two measures visually we also see, that the big gap between the top two teams and the other teams suggested by the OPR ranking, turns into a group of four teams with a gap to the remaining teams:

```{r scatter, message = FALSE}
#| label: fig-scatter
#| fig.cap: Scatterplot of team contributions measured in OPR (x axis) and to the 
#|   score without foul points (y axis). 
#| fig.alt: >
#|   Scatterplot of team contributions measured in OPR (x axis) and to the 
#|   score without foul points (y axis). 
#|   The relationship is roughly linear, but the spacing between team contributions
#|   is different as discussed in more detail in the writeup.
library(tidyverse)
library(ggplot2)

# combine the OPR with the ranking not including fouls:
all_points <- scores %>% get_ranking_by(alliances_score, alliances_score-score_breakdown_foulPoints)
all_points %>% ggplot(aes(x = `rating(alliances_score)`, y = `rating(alliances_score - score_breakdown_foulPoints)`)) + 
  geom_point() + xlab("OPR") + 
  ggrepel::geom_label_repel(aes(label = team_key), size=3, alpha = 0.6) +
  geom_rug(length = unit(0.03, "npc")) +
  scale_y_continuous(expand = c(0.2, 0.2)) +
  scale_x_continuous(expand = c(0.2, 0.2)) +
  coord_equal()
```

### Worst fouling team

Any of the details going into the score can be used as outcome measurement.
The number of points given to the opponent because of a team's foul is determined (in the 2024 season) as 5 points for a foul and 2 points for a technical foul. The list below is sorted from worst offenders to least offenders. A negative number might be interpreted as a team's ability to provoke fouls from the opponents.

```{r}
scores %>% filter(comp_level=="qm") %>%
  get_ranking_by(2*score_breakdown_foulCount+5*score_breakdown_techFoulCount) 
```

### Best at categories

```{r}
names(scores) <- gsub("score_breakdown_", "", names(scores))
detailed <- scores %>% filter(comp_level=="qm") %>%
  get_ranking_by(alliances_score, alliances_score - foulPoints, autoPoints, teleopPoints, -1*(2*foulCount+5*techFoulCount)) 

# add the tba rating to the mix:
tba_rating <- tba_ranking("2024iawes")
detailed <- detailed %>% left_join(tba_rating %>% select(team_key, TBA=rank), by="team_key")
```

Taking more measures into account, we can see in the parallel coordinate plot below, that the third place TBA ranking for team frc9999 is not backed up by a similar high performance in the contribution to the score, which is indicative of a lot of luck and/or strategy in the qualifying matches. On the other hand, team frc4646 is in the top three performing teams except for their TBA ranking, indicating that they would be a great pick for any alliance. Similarly, team frc525 would make a good contributing partner in an alliance. 

```{r parallel}
#| label: fig-parallel
#| fig.cap: Parallel Coordinate Plot of multiple measures showing different aspects of a team's contribution. Higher is better for all measures.
#| fig.alt: Parallel Coordinate Plot of multiple measures showing different aspects of a team's contribution. From left to right, the teams' tba ranking and their contribution to the score, the score without fouls, the autoPoints, the teleop points, and the number of fouls are shown. Team frc3928 dominates in all of these measures except for the number of auto points and the number of fouls.

library(ggpcp)


 detailed %>% 
   mutate(
     TBA = max(TBA)-TBA,
    team_key = reorder(factor(team_key), TBA) 
    ) %>%
   pcp_select(team_key, TBA, starts_with("rating"), team_key) %>%
   pcp_scale() %>%
   ggplot(aes_pcp()) + 
   geom_pcp() + 
   geom_pcp_labels(size=3) + 
   xlab("") + ylab("") + 
   theme(axis.text.x = element_text(angle=30, hjust=1))

```

