---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tba

<!-- badges: start -->
<!-- badges: end -->

The goal of the tba package is to provide easy access to match data for events 
run in the first robotics division. Access is based on the api of [The Blue Alliance](https://www.thebluealliance.com/). 

## Installation

You can install the development version of tba from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("heike/tba")
```

In order to access the API of The Blue Alliance, you will need to [sign up for a (free) account](https://www.thebluealliance.com/apidocs/v3). 
Then run 
``` r
tba::set_api_key(key = "jnn1ZiUfqPuA9UJyV9EVa3jtL5qudBBlYXKPEVQDeQgQgz5L8BypwjTzVKxtCs3W") 
# key is not valid, get your own!
```


## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(tba)
## basic example code

# Team Neutrino has team code `frc3928` 
get_records("team/frc3928")

# In the 2024 season, the team attended the following events:
events <- get_records("team/frc3928/events/2024")
events$name
```
The matches from Clash in the Corn can be accessed using the event code `2024iawes`:

```{r}
library(dplyr, quietly=TRUE)
matches <- get_records("event/2024iawes/matches")
count(matches, comp_level)
# 22 qualifying matches
# 5 semi-finals
# 2 finals
```

The function `get_match_details` turns the records into a more manageable form by turning the match-based rows from the `get_records` results into team-based results, i.e.  every row corresponds to the 
detailed scores of each team in each one of the matches:

```{r}
scores <- get_match_details(matches)
head(scores %>% select(comp_level, match_number, set_number, alliance, team_key, score, ends_with("Points")))
```